### 2015-07-10 narrative/intro.tex
reading perspective.
work in progress.

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing, storing and retrieving the large volumes of data that are 
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing and retrieving  data
where the user and the storage system work together to optimize performance with
respect to user requirements, current system state, and characteristics of the tiered-storage system.
Our investigation will include  novel techniques for the efficient mapping of data objects
from the user space across to a hierarchical tiered storage layer. These techniques will
include support for application-guided data reductions with user-defined error bounds
and transformations to optimize performance, capacity and minimize bandwidth bottlenecks.

% 
Our proposed research program is guided by the following overarching objectives:
\begin{tightItemize}
\item
{\em Address emerging storage architecture and technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best''  data, which is automatically placed throughout the
entire storage stack.
\item
{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
supported data lifetime, our approach will enable researchers to 
place more knowledge into the Storage Systems and Input/Output (SSIO) system so that data can be recast from 
a sequence of bytes to information rich data. This information rich data can then
be used by the system to guide the  placement within the tiered storage and will optimize the time
to gain knowledge and insight from exascale simulations.

@@

\item
{\em Emphasize impacts on science:}
%The metric we are most interested in optimizing is time to knowledge.  
%Current approaches to addressing the I/O bottleneck fall into two broad categories:
%parallel file system approaches that optimize the throughput for an entire
%system, and I/O middleware approaches that optimize the performance of a single
%application. Both approaches have been successful to date, but are unlikely to
%overcome the remaining obstacles in reaching exascale. 
Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
As extracting knowledge is the only metric that matters for scientific applications, we will
focus not only upon data storage, but upon efficient  retrieval of knowledge from their 
workflows.
Moreover, we aim to perform the optimization not only for a single
application, but rather for multiple workloads running on a multi-user, multi-application
system.
\end{tightItemize}

This proposal can be summarized as research into using the varying importance of simulation data over time to
ensure that the right portions of data, in the right quality  are available to scientists when they are needed.
Application-specific information about the importance of data will be used in conjunction with system and resource-specific information
to optimize  scientists experience in interacting with, and extracting knowledge and insight from their  data.

@@

This proposal combines knowledge and expertise 
from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
management and storage systems knowledge of UCSC, and the middleware expertise
of Rutgers. We will integrate this diverse expertise to investigate how to create
a data, workflow, and application aware storage and I/O system.
%This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
%file system (SNL, UCSC) and storage (UCSC), and connects and
%coordinates these key storage components in a seamless fashion.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline

>>>>

Avoiding impending bottlenecks in exascale scientific discovery will require
new research into managing, storing and retrieving the large volumes of data that are 
produced by simulations and analyzed for months afterwards.
%
In this project we will demonstrate a cooperative approach for storing and retrieving  data
where the user and the storage system work together to optimize performance with
respect to user requirements, current system state, and characteristics of the tiered-storage system.
Our investigation will include  novel techniques for the efficient mapping of data objects
from the user space to a hierarchical tiered storage layer. These techniques will
include support for application-guided data reductions with user-defined error bounds
and transformations to optimize performance, capacity and minimize bandwidth bottlenecks.

% 
Our proposed research program is guided by the following overarching objectives:
\begin{tightItemize}
\item
{\em Address emerging storage architecture and technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best''  data, which is automatically and appropriately placed within the
 storage stack.
\item
{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
supported data lifetime, our approach will enable researchers to 
place more knowledge into the Storage Systems and Input/Output (SSIO) system 
so that data can be recast from a sequence of bytes to information rich data. This 
information rich data can then be used by the system to guide the  placement within 
the tiered storage and will optimize the time to gain knowledge and insight from 
exascale simulations.

@@

\item
{\em Emphasize impacts on science:}
%The metric we are most interested in optimizing is time to knowledge.  
%Current approaches to addressing the I/O bottleneck fall into two broad categories:
%parallel file system approaches that optimize the throughput for an entire
%system, and I/O middleware approaches that optimize the performance of a single
%application. Both approaches have been successful to date, but are unlikely to
%overcome the remaining obstacles in reaching exascale. 
Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
As extracting knowledge is the only metric that matters for scientific applications, we will
focus not only on data storage, but also on efficient retrieval of knowledge from their 
workflows.
Moreover, we aim to perform the optimization not only for a single
application, but rather for multiple workloads running on a multi-user, multi-application
system.
\end{tightItemize}

%%% Manish: I am not sure I fully understand the first part of the paragraph below.
This proposal can be summarized as research into using the varying importance of simulation data over time to
ensure that the right portions of data, with the right quality  are available to scientists when they are needed.
Application-specific information about the importance of data will be used in conjunction with system and resource-specific information
to optimize the scientists' experience in interacting with, and extracting knowledge and insight from their data.

@@

This proposal combines knowledge and expertise 
from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
management and storage systems knowledge at UCSC, and the middleware expertise
at Rutgers. We will integrate this diverse expertise to investigate how to create
a data, workflow, and application aware storage and I/O system.
%This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
%file system (SNL, UCSC) and storage (UCSC), and connects and
%coordinates these key storage components in a seamless fashion.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
